---
title: "LRs and Clustering"
author: "Kiegan Rice"
output: html_document
---

```{r}
library(tidyverse)
library(gridExtra)
library(matrixStats)

train <- read_csv("../data/train.csv")
test <- read_csv("../data/test.csv")
all <- full_join(train, test)

```


let's grab columns: 16, 33, 43, 65, 73, 80, 82, 90, 91, 101, 108, 117, 133, 134, 165, 183, 189, 194, 199, 217, 226, 237, 258, 295, 298
```{r}

lasso_cols_1se <- c(16, 33, 43,
                65,73, 80, 82, 90, 91, 
                101, 108, 117, 133, 134, 165, 183, 189, 194, 199, 
                217, 226, 237, 258, 295, 298)

train_cols_1se <- as.data.frame(train[,c(1:2, lasso_cols_1se+3)])
test_cols_1se <- as.data.frame(test[,c(1, lasso_cols_1se+2)])


names(train_cols_1se) <- paste0("col_", names(train_cols_1se))
names(test_cols_1se) <- paste0("col_", names(test_cols_1se))
```

Functions to calculate kernel density estimates for LRs...

```{r}
density_lr <- function(dataset, col_name){
  class0 <- dataset[dataset$col_target == 0,]
  class1 <- dataset[dataset$col_target == 1,]
  
  class0_density <- density(class0[,col_name])
  class1_density <- density(class1[,col_name])
  
  lr_density <- class1_density$y/class0_density$y
  
  new_col <- paste0("kernel_lr_", col_name)
  
  for(i in 1:nrow(dataset)){
    dataset[i,new_col] <- lr_density[which.min(abs(dataset[i, col_name] - class0_density$x))]
  }
  return(dataset)
}

density_lr_est <- function(dataset, col_name){
  class0 <- dataset[dataset$col_target == 0,]
  class1 <- dataset[dataset$col_target == 1,]
  
  class0_density <- density(class0[,col_name])
  class1_density <- density(class1[,col_name])
  
  lr_density <- class1_density$y/class0_density$y
  
  return(data.frame(density_x = class0_density$x, density_lr = lr_density))
}

test_density_cols <- function(dataset_test, col_name, density_df){
  new_col <- paste0("kernel_lr_", col_name)
  for(i in 1:nrow(dataset_test)){
    dataset_test[i,new_col] <- density_df$density_lr[which.min(abs(dataset_test[i,col_name] - density_df$density_x))]
  }
  return(dataset_test)
}

```

Calculate kernel density LR features for each of the columns that remained in our 1se LASSO model. 

```{r}
train_cols_1se <- density_lr(train_cols_1se, "col_16")
train_cols_1se <- density_lr(train_cols_1se, "col_33")
train_cols_1se <- density_lr(train_cols_1se, "col_43")
train_cols_1se <- density_lr(train_cols_1se, "col_65")
train_cols_1se <- density_lr(train_cols_1se, "col_73")
train_cols_1se <- density_lr(train_cols_1se, "col_80")
train_cols_1se <- density_lr(train_cols_1se, "col_82")
train_cols_1se <- density_lr(train_cols_1se, "col_90")
train_cols_1se <- density_lr(train_cols_1se, "col_91")
train_cols_1se <- density_lr(train_cols_1se, "col_101")
train_cols_1se <- density_lr(train_cols_1se, "col_108")
train_cols_1se <- density_lr(train_cols_1se, "col_117")
train_cols_1se <- density_lr(train_cols_1se, "col_133")
train_cols_1se <- density_lr(train_cols_1se, "col_134")
train_cols_1se <- density_lr(train_cols_1se, "col_165")
train_cols_1se <- density_lr(train_cols_1se, "col_183")
train_cols_1se <- density_lr(train_cols_1se, "col_189")
train_cols_1se <- density_lr(train_cols_1se, "col_194")
train_cols_1se <- density_lr(train_cols_1se, "col_199")
train_cols_1se <- density_lr(train_cols_1se, "col_217")
train_cols_1se <- density_lr(train_cols_1se, "col_226")
train_cols_1se <- density_lr(train_cols_1se, "col_237")
train_cols_1se <- density_lr(train_cols_1se, "col_258")
train_cols_1se <- density_lr(train_cols_1se, "col_295")
train_cols_1se <- density_lr(train_cols_1se, "col_298")



test_cols_1se <- test_density_cols(test_cols_1se, "col_16", density_df = density_lr_est(train_cols_1se, "col_16"))
test_cols_1se <- test_density_cols(test_cols_1se, "col_33", density_df = density_lr_est(train_cols_1se, "col_33"))
test_cols_1se <- test_density_cols(test_cols_1se, "col_43", density_df = density_lr_est(train_cols_1se, "col_43"))
test_cols_1se <- test_density_cols(test_cols_1se, "col_65", density_df = density_lr_est(train_cols_1se, "col_65"))
test_cols_1se <- test_density_cols(test_cols_1se, "col_73", density_df = density_lr_est(train_cols_1se, "col_73"))
test_cols_1se <- test_density_cols(test_cols_1se, "col_80", density_df = density_lr_est(train_cols_1se, "col_80"))
test_cols_1se <- test_density_cols(test_cols_1se, "col_82", density_df = density_lr_est(train_cols_1se, "col_82"))
test_cols_1se <- test_density_cols(test_cols_1se, "col_90", density_df = density_lr_est(train_cols_1se, "col_90"))
test_cols_1se <- test_density_cols(test_cols_1se, "col_91", density_df = density_lr_est(train_cols_1se, "col_91"))
test_cols_1se <- test_density_cols(test_cols_1se, "col_101", density_df = density_lr_est(train_cols_1se, "col_101"))
test_cols_1se <- test_density_cols(test_cols_1se, "col_108", density_df = density_lr_est(train_cols_1se, "col_108"))
test_cols_1se <- test_density_cols(test_cols_1se, "col_117", density_df = density_lr_est(train_cols_1se, "col_117"))
test_cols_1se <- test_density_cols(test_cols_1se, "col_133", density_df = density_lr_est(train_cols_1se, "col_133"))
test_cols_1se <- test_density_cols(test_cols_1se, "col_134", density_df = density_lr_est(train_cols_1se, "col_134"))
test_cols_1se <- test_density_cols(test_cols_1se, "col_165", density_df = density_lr_est(train_cols_1se, "col_165"))
test_cols_1se <- test_density_cols(test_cols_1se, "col_183", density_df = density_lr_est(train_cols_1se, "col_183"))
test_cols_1se <- test_density_cols(test_cols_1se, "col_189", density_df = density_lr_est(train_cols_1se, "col_189"))
test_cols_1se <- test_density_cols(test_cols_1se, "col_194", density_df = density_lr_est(train_cols_1se, "col_194"))
test_cols_1se <- test_density_cols(test_cols_1se, "col_199", density_df = density_lr_est(train_cols_1se, "col_199"))
test_cols_1se <- test_density_cols(test_cols_1se, "col_217", density_df = density_lr_est(train_cols_1se, "col_217"))
test_cols_1se <- test_density_cols(test_cols_1se, "col_226", density_df = density_lr_est(train_cols_1se, "col_226"))
test_cols_1se <- test_density_cols(test_cols_1se, "col_237", density_df = density_lr_est(train_cols_1se, "col_237"))
test_cols_1se <- test_density_cols(test_cols_1se, "col_258", density_df = density_lr_est(train_cols_1se, "col_258"))
test_cols_1se <- test_density_cols(test_cols_1se, "col_295", density_df = density_lr_est(train_cols_1se, "col_295"))
test_cols_1se <- test_density_cols(test_cols_1se, "col_298", density_df = density_lr_est(train_cols_1se, "col_298"))

```


Let's try a LASSO and see what happens...

```{r}
lasso_kernel_lrs <- cv.glmnet(x = as.matrix(train_cols_1se[,-c(1:2)]), y = train$target, family = "binomial", alpha = 1, nfolds = 10)
s = lasso_kernel_lrs$lambda.min
s_1se = lasso_kernel_lrs$lambda.1se
lasso_kernel_coef = predict(lasso_kernel_lrs, s = s, type = "coefficients")
lasso_kernel_coef
```

Now let's try some model-based clustering...

```{r}
library(mclust)
clPairs(data = train_cols_1se[,c(28:31, 33:34, 36:38, 43:44, 46, 48:50, 52)], classification = train_cols_1se$col_target)


BIC <- mclustBIC(data = train_cols_1se[,c(28:31, 33:34, 36:38, 43:44, 46, 48:50, 52)])
plot(BIC)
summary(BIC)


mod1 <- Mclust(train_cols_1se[,c(28:31, 33:34, 36:38, 43:44, 46, 48:50, 52)], x = BIC)
summary(mod1, parameters = TRUE)
plot(mod1, what = "classification")

table(mod1$classification, train_cols_1se$col_target)
```


```{r}
test_cols_1se$col_target = -1
all_1se <- full_join(train_cols_1se, test_cols_1se)
head(all_1se)
all_1se <- all_1se %>% select(-target)

clPairs(data = all_1se[,c(28:31, 33:34, 36:38, 43:44, 46, 48:50, 52)], classification = all_1se$col_target)
BIC <- mclustBIC(data = all_1se[,c(28:31, 33:34, 36:38, 43:44, 46, 48:50, 52)])
plot(BIC)
summary(BIC)

mod1 <- Mclust(all_1se[,c(28:31, 33:34, 36:38, 43:44, 46, 48:50, 52)], x = BIC)
summary(mod1, parameters = TRUE)
table(mod1$classification, all_1se$col_target)
plot(mod1, what = "classification")

all_1se$cluster_class <- as.factor(mod1$classification)

all_1se$in_cluster9 <- ifelse(mod1$classification == 9, 1, 0)
all_1se$in_cluster4or5 <- ifelse(mod1$classification == 4 | mod1$classification == 4, 1, 0)
all_1se$in_cluster1 <- ifelse(mod1$classification == 1, 1, 0)


train_1se <- all_1se %>% filter(col_target != -1)
test_1se <- all_1se %>% filter(col_target == -1)
```


Now that we have some clustering, let's see if including that in our LASSO model helps at all. 

```{r}
lasso_lrs_clusters <- cv.glmnet(x = as.matrix(train_1se[,-c(1:2, 53)]), y = train_1se$col_target, family = "binomial", alpha = 1, nfolds = 10)
s = lasso_lrs_clusters$lambda.min
s_1se = lasso_lrs_clusters$lambda.1se
lasso_clusters_coef = predict(lasso_lrs_clusters, s = s_1se, type = "coefficients")
lasso_clusters_coef
```

Instead of LASSO, let's try a random forest instead.

```{r}
library(randomForest)
library(caret)
#trControl <- trainControl(method = "cv", number = 10)

#first_rf <- caret::train(form = factor(col_target)~., data = train_1se[,-c(1, 53)], 
#                       method = "rf",
#                       trControl = trControl)

first_rf <- randomForest(factor(col_target)~., data=train_1se[,-c(1)])
preds_train_rf <- predict(first_rf, newdata = train_1se[,-c(1:2)], type = "prob")
test_rf <- data.frame(id = train$id, target = train$target, preds = as.vector(preds_train_rf[,2]))
preds_rf <- predict(first_rf, newdata = test_1se[,-c(1:2)], type = "prob")
test <- data.frame(id = test$id, preds = preds_rf[,2])
```




```{r}
preds <- predict(lrs_clusters_model, s = s, newx = as.matrix(test_1se[,-c(1:2, 53)]), type = "response") 

preds <- predict(lrs_clusters_model, s = s, newx = as.matrix(train_1se[,-c(1:2, 53)]), type = "response")

test <- data.frame(id = train$id, target = train$target, preds = as.vector(preds$fit))

submission <- data.frame(id = test$id, target = preds_rf[,2])
write_csv(submission, "lasso_lrs_clusters.csv")
```

